Exploitation Techniques and Mitigations

### 2016-01-31 docs/paper.tex
Paper: Fix some typos

Exploiting binaries is an extremely broad topic with many specialized
techniques for even the most exotic scenarios. In this writeup we will take a
narrow look at exploiting \texttt{printf} with crafted format strings first and
follow up with an introduction about buffer overflows. Mitigation mechanisms
will be disabled at first and enabled one by one --- discussing them when they
are put into place. The buffer overflow will be augmented to inject and execute
shell code which is then prevented by the Data Execution Prevention (DEP)
mechanism.

@@

Although x86 has been chosen as target platform a quick glance at other
architectures (x86\_64 and ARM) is taken prior concluding this writeup. Some
basics about the target platform will be communicated before running the first
exploit.

@@

Exploiting binaries was comparatively easy ten to fifteen years ago. There were
no special mitigation mechanisms in place denying even the most simplest
exploits. This is the point in time where we will start off. First we talk
about two very simple exploits, namely the format string exploit and the buffer
overflow in combination with shell code. Although there is a huge collection of
exploitation techniques known to the public, we will only look at a very small
fraction of them in this project.

@@

Examples will be provided along the way to support the reader and provide
additional explanations. Finally we will conclude with a word about other
architectures (x86\_64 and ARM) followed by a short recap about what has been
taught in this writeup.

@@

The second type of exploits we'll look at is known as buffer overflows and as
on may already derive from the name, this is about submitting more data to a
buffer than it was originally designed for. This setup can be exploited when
bound checking is done wrong or not at all. An attacker is therefore able to
overwrite memory behind the buffer's location.

@@

If an attacker can provide the data written to the buffer and no (or wrong)
bound checking is done, he is able to inject arbitrary (malicious) code into
the stack frame. This could, for instance, be used to overwrite a flag
indication whether an authentication has been performed successfully or not.
But since this is pretty forward let's go beyond that and see what happens when
changing the return address.

@@

A more com comprehensive article \cite{phrack_stack_smash} about Stack Smashing
is available on \texttt{phrack}\footnote{\url{http://phrack.org/}}.

@@

The final \texttt{ret} statement of a gadgets is required for chaining them
together. We not only use the buffer overflow to control \emph{one}, but
\emph{multiple} return addresses. Because of this our payload will be composed
of a list of return addresses (the start of each gadgets) interleaved by some
padding. Each provided address will be consumed one by one at the end of each
gadgets to \emph{jump} to the next one. The payload may still contain data like
\texttt{"/bin//sh"} if necessary.

@@

One can easily get a list of available gadgets by piping the output of
\texttt{objdump} to \texttt{grep} filtering for \texttt{ret} instructions. This
is done in \cref{src:rop_gadgets} where three different gadgets can be
observed. Of course each gadget can be arbitrary long, we just used a length of
5 instructions in this example.

@@

Breaking ASLR, even when the code is compiled with \texttt{-pie}, is easier
than it seems at first. Relocation only happens to a section as a whole,
functions inside a section still share the same relative distance as they would
without ASLR. But before exploit this fact, have a look at the randomized
addresses again.

@@

    We now know that \texttt{system} is at:
    \[ \mathtt{0xb7e72280} - \mathtt{0xd0f0} = \mathtt{0xb7e65190} \]
\end{quote}

@@

CFI is a big topic and, like other topics already mentioned, goes beyond the
scope of this writeup. The first pointer we hand the reader aims at the
corresponding
section\footnote{\url{http://clang.llvm.org/docs/ControlFlowIntegrity.html}} of
the Clang documentation, but we recommend checking out the related research
paper \cite{ms_cfi} for more information. A more accessible and recent way to
this topic may be the
talk\footnote{\url{https://www.youtube.com/watch?v=FA0VK7s5tSQ}} \textit{New
memory corruption attacks: why can't we have nice things?} given by gannimo
(Mathias Payer).

@@

Starting with no mitigation mechanisms in place, we have seen how easy it is to
manipulate the program by exploiting just one simple buffer overflow. Going
beyond simple manipulations like changing locale variables, we craft shell code
and injected it to open up a shell accepting and executing arbitrary inputs.
Next, Data Execution Prevention (DEP) was presented to deny the ability of
\emph{injecting} new code.

@@

Address Space Layout Randomization (ASLR) was established to prevent ROP but
can be defeated with an information leak. Even the StackGuard mechanism can be
broking with brute force (including an unexpected low upper bound) in certain
scenarios.

>>>>

Exploiting binaries is an extremely broad topic with many specialized
techniques for even the most exotic scenarios. In this write-up we will take a
narrow look at exploiting \texttt{printf} with crafted format strings first and
follow up with an introduction about buffer overflows. Mitigation mechanisms
will be disabled at first and enabled one by one --- discussing them when they
are put into place. The buffer overflow will be augmented to inject and execute
shell code which is then prevented by the Data Execution Prevention (DEP)
mechanism.

@@

Although x86 has been chosen as target platform a quick glance at other
architectures (x86\_64 and ARM) is taken prior concluding this write-up. Some
basics about the target platform will be communicated before running the first
exploit.

@@

Exploiting binaries was comparatively easy ten to fifteen years ago. There were
no special mitigation mechanisms in place denying even the simplest exploits.
This is the point in time where we will start off. First we talk about two very
simple exploits, namely the format string exploit and the buffer overflow in
combination with shell code. Although there is a huge collection of
exploitation techniques known to the public, we will only look at a very small
fraction of them in this project.

@@

Examples will be provided along the way to support the reader and provide
additional explanations. Finally we will conclude with a word about other
architectures (x86\_64 and ARM) followed by a short recap about what has been
taught in this write-up.

@@

The second type of exploits we'll look at is known as buffer overflows and as
one may already derive from the name, this is about submitting more data to a
buffer than it was originally designed for. This setup can be exploited when
bound checking is done wrong or not at all. An attacker is therefore able to
overwrite memory behind the buffer's location.

@@

If an attacker can provide the data written to the buffer and no (or wrong)
bound checking is done, he is able to inject arbitrary (malicious) code into
the stack frame. This could, for instance, be used to overwrite a flag
indication whether an authentication has been performed successfully or not.
But since this is pretty straight forward let's go beyond that and see what
happens when changing the return address.

@@

A more comprehensive article \cite{phrack_stack_smash} about Stack Smashing is
available on \texttt{phrack}\footnote{\url{http://phrack.org/}}.

@@

The final \texttt{ret} statement of a gadget is required for chaining them
together. We not only use the buffer overflow to control \emph{one}, but
\emph{multiple} return addresses. Because of this our payload will be composed
of a list of return addresses (the start of each gadgets) interleaved by some
padding. Each provided address will be consumed one by one at the end of each
gadgets to \emph{jump} to the next one. The payload may still contain data like
\texttt{"/bin//sh"} if necessary.

@@

One can easily get a list of available gadgets by piping the output of
\texttt{objdump} to \texttt{grep} filtering for \texttt{ret} instructions. This
is done in \cref{src:rop_gadgets} where three different gadgets can be
observed. Of course each gadget can be arbitrarily long, we just used a length
of 5 instructions in this example.

@@

Breaking ASLR, even when the code is compiled with \texttt{-pie}, is easier
than it seems at first. Relocation only happens to a section as a whole,
functions inside a section still share the same relative distance as they would
without ASLR. But before exploiting this fact, have a look at the randomized
addresses again.

@@

    We now know that \texttt{system} is at:
    \[ \texttt{0xb7e72280} - \texttt{0xd0f0} = \texttt{0xb7e65190} \]
\end{quote}

@@

CFI is a big topic and, like other topics already mentioned, goes beyond the
scope of this write-up. The first pointer we hand the reader aims at the
corresponding
section\footnote{\url{http://clang.llvm.org/docs/ControlFlowIntegrity.html}} of
the Clang documentation, but we recommend checking out the related research
paper \cite{ms_cfi} for more information. A more accessible and recent way to
this topic may be the
talk\footnote{\url{https://www.youtube.com/watch?v=FA0VK7s5tSQ}} \textit{New
memory corruption attacks: why can't we have nice things?} given by gannimo
(Mathias Payer).

@@

Starting with no mitigation mechanisms in place, we have seen how easy it is to
manipulate the program by exploiting just one simple buffer overflow. Going
beyond simple manipulations like changing local variables, we craft shell code
and injected it to open up a shell accepting and executing arbitrary inputs.
Next, Data Execution Prevention (DEP) was presented to deny the ability of
\emph{injecting} new code.

@@

Address Space Layout Randomization (ASLR) was established to prevent ROP but
can be defeated with an information leak. Even the StackGuard mechanism can be
broken with brute force (including an unexpected low upper bound) in certain
scenarios.
